---
title: "CLSA COVID-19 Epigenetic Analysis"
author: "Ryan Stanley Falck"
date: "06/07/2022"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Data Management

```{r}
#Load variables and packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,Hmisc, knitr, pander, broom, haven, rms, nlme,lspline, lavaan, lme4, semTools, tableone, psych, ggpubr, ggplot2, openxlsx, expss, effects,
               BiocManager, minfi, IlluminaHumanMethylationEPICmanifest, IlluminaHumanMethylationEPICanno.ilm10b4.hg19, remotes, ewastools)
setwd("~/Desktop/UBC-Postdoctoral Fellowship/CLSA - Covid Analyses")
COVID<-read.xlsx("CLSA_COVID_CKB_v2.xlsx")#Processed Baseline + Follow-up 1 + COVID Questionnaire

#Check epigenetic variables
describe(COVID$DNAmAge_COM_0)
describe(COVID$Hannum_Age_COM_0)
describe(COVID$AgeAccelerationDifference_COM_0)
describe(COVID$AgeAccelerationResidual_COM_0)

#Exclude people without Epigenetic data
COVID.1<-subset(COVID, DNAmAge_COM_0>=0) #879 participants with Epigenetic data in Comprehensive Cohort who also completed COVID-19 Questionnaire Study

# Excluded individuals with dementia, subjective cognitive impairment,
# anxiety, depression, or mood disorders

COVID.Final <- subset(COVID.1, Dementia_0 == "No" & SCI_0 == "No" & Anxiety_0 == "No" &
                        Clin_Depression_0 == "No" & Mood_Disord_0 == "No") # 663 participants included in the study


```

# Step 2: Baseline Characteristics
```{r}
BL_vars<-dput(names(COVID.Final[c(7,6,9,10,8,11,12,13,34,35,36,42,2:5)]))

Table1_continuous<-CreateTableOne(vars=BL_vars, data=COVID.Final)
print(Table1_continuous,contDigits=2,missing=TRUE,quote=TRUE)
```

# Step 3: Predicting Depressive Symptoms from Epigenetic Age

## 3.1: Linear Mixed Model Set-up
```{r}
library(dplyr)

COVID.Final.BL<- COVID.Final[c(1:13,34,35,42,36)]
COVID.Final.FU1<- COVID.Final[c(1:13,34,35,42,64)]
COVID.Final.CBL<- COVID.Final[c(1:13,34,35,42,68)]
COVID.Final.CEX<- COVID.Final[c(1:13,34,35,42,73)]

COVID.Final.BL$time <- 1
COVID.Final.FU1$time <- 2
COVID.Final.CBL$time <- 3
COVID.Final.CEX$time <- 4

COVID.Final.BL<- rename(COVID.Final.BL, c("CESD_10_0" = "CESD_10"))
COVID.Final.FU1<- rename(COVID.Final.FU1, c("CESD_10_1" = "CESD_10"))
COVID.Final.CBL<- rename(COVID.Final.CBL, c("CESD_10_Baseline"="CESD_10"))
COVID.Final.CEX<- rename(COVID.Final.CEX, c("CESD_10_Exit"="CESD_10"))

COVID_Full_Complete<- rbind(COVID.Final.BL,COVID.Final.FU1,COVID.Final.CBL,COVID.Final.CEX)

COVID_Full_Complete$timefactor <- as.factor(COVID_Full_Complete$time)
```



## 3.2: DNA Methylation Age


### 3.2.1: Crude Model
```{r}
DNA_CESD_model_crude<-lmer(CESD_10~DNAmAge_COM_0*timefactor + (1 + time|ID), REML = TRUE, COVID_Full_Complete)
summary(DNA_CESD_model_crude)
anova(DNA_CESD_model_crude)#significant DNA methylation x time interaction
```


### 3.2.2: Full Model - Adjusted for age, sex, income, education, living status, alcohol intake, smoking status, physical activity level, & BMI
```{r}
DNA_CESD_model_full<-lmer(CESD_10~DNAmAge_COM_0*timefactor + (1 + time|ID) + Age + Sex + Income_Level + Education4 + Living_Status + 
                            Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_Full_Complete)
summary(DNA_CESD_model_full)
anova(DNA_CESD_model_full)#significant  DNA methylation x time interaction
```


### 3.2.3: Simple Slopes Analyses (Contrasts)
```{r}
mean(COVID.Final$DNAmAge_COM_0) #Mean DNA Methylation Age = 57.49334
mean(COVID.Final$DNAmAge_COM_0) + 2*sd(COVID.Final$DNAmAge_COM_0) #High" DNA Methylation Age = 74.9256 
mean(COVID.Final$DNAmAge_COM_0) - 2*sd(COVID.Final$DNAmAge_COM_0) #"Low" DNA Methylation Age = 40.06108


ef1<-effect(term = "DNAmAge_COM_0*timefactor", xlevels=list(DNAmAge_COM_0=c(40.06108,57.49334,74.9256)), mod=DNA_CESD_model_full)
efdata1<-as.data.frame(ef1)
efdata2<-efdata1[c(2,3,4,1)]

efdata2$DNA<-NA
efdata2$DNA[efdata2$DNAmAge_COM_0 == 40.06108] <- "Low"
efdata2$DNA[efdata2$DNAmAge_COM_0 == 57.49334] <- "Mean"
efdata2$DNA[efdata2$DNAmAge_COM_0 == 74.9256] <- "High"

efdata2
#Contrast 1: Differences between High and Low DNA from BL to FU1
Diff1= efdata2$fit[c(3)] - efdata2$fit[c(6)] 
Diff2= efdata2$fit[c(1)] - efdata2$fit[c(4)]
Diff3= Diff1 - Diff2
se1 = sqrt((efdata2$se[c(3)])^2 + (efdata2$se[c(6)])^2)
se2 = sqrt((efdata2$se[c(1)])^2 + (efdata2$se[c(4)])^2)
se3 = sqrt(se1^2 + se2^2)
t1 = Diff3/se3
n=663
p=pt(-abs(t1),df=n-1)
LL3= Diff3 - 1.64*se3
UL3=  Diff3 + 1.64*se3
Diff1
Diff2
Diff3
se1
se2
se3
t1
p
LL3
UL3

#Contrast 2: Differences between High and Low DNA from FU1 to COVID BL
Diff1.2= efdata2$fit[c(6)] - efdata2$fit[c(9)] 
Diff2.2= efdata2$fit[c(4)] - efdata2$fit[c(7)]
Diff3.2= Diff1.2 - Diff2.2
se1.2 = sqrt((efdata2$se[c(6)])^2 + (efdata2$se[c(9)])^2)
se2.2 = sqrt((efdata2$se[c(4)])^2 + (efdata2$se[c(7)])^2)
se3.2 = sqrt(se1.2^2 + se2.2^2)
t2 = Diff3.2/se3.2
p.2=pt(-abs(t2),df=n-1)
LL3.2= Diff3.2 - 1.64*se3.2
UL3.2=  Diff3.2 + 1.64*se3.2
Diff1.2
Diff2.2
Diff3.2
se1.2
se2.2
se3.2
t2
p.2
LL3.2
UL3.2

#Contrast 3: Differences between High and Low DNA from COVID BL to COVID Exit
Diff1.3= efdata2$fit[c(9)] - efdata2$fit[c(12)] 
Diff2.3= efdata2$fit[c(7)] - efdata2$fit[c(10)]
Diff3.3= Diff1.3 - Diff2.3
se1.3 = sqrt((efdata2$se[c(6)])^2 + (efdata2$se[c(9)])^2)
se2.3 = sqrt((efdata2$se[c(4)])^2 + (efdata2$se[c(7)])^2)
se3.3 = sqrt(se1.3^2 + se2.3^2)
t3 = Diff3.3/se3.3
p.3=pt(-abs(t3),df=n-1)
LL3.3= Diff3.3 - 1.64*se3.3
UL3.3=  Diff3.3 + 1.64*se3.3
Diff1.3
Diff2.3
Diff3.3
se1.3
se2.3
se3.3
t3
p.3
LL3.3
UL3.3
```


### 3.2.4: Graphing Simple Slopes
```{r}
efdata1$DNA<-NA
efdata1$DNA[efdata1$DNAmAge_COM_0==40.06108]<-"-2 SD"
efdata1$DNA[efdata1$DNAmAge_COM_0==57.49334]<-"Mean"
efdata1$DNA[efdata1$DNAmAge_COM_0==74.9256]<-"+2 SD"

efdata1$Timepoint<-NA
efdata1$Timepoint[efdata1$timefactor==1]<-"Baseline"
efdata1$Timepoint[efdata1$timefactor==2]<-"First Follow-up"
efdata1$Timepoint[efdata1$timefactor==3]<-"COVID-19 Study Entry"
efdata1$Timepoint[efdata1$timefactor==4]<-"COVID-19 Study Exit"
efdata1$Timepoint<-ordered(efdata1$Timepoint, levels = c("Baseline", "First Follow-up", "COVID-19 Study Entry", "COVID-19 Study Exit"))

ggplot(data=efdata1, aes(x=Timepoint, y=fit, group=DNA)) + geom_line(aes(color=DNA)) + ylim(0,8) + 
  labs(y= "CESD-10 Score", color="DNA Methylation") + theme_classic() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                                                                              panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```



## 3.3: Hannum Age


### 3.3.1: Crude Model
```{r}
Hannum_CESD_model_crude<-lmer(CESD_10~Hannum_Age_COM_0*timefactor + (1 + time|ID), REML = TRUE, COVID_Full_Complete)
summary(Hannum_CESD_model_crude)
anova(Hannum_CESD_model_crude)#significant Hannum Age x time interaction
```


### 3.3.2: Full Model - Adjusted for age, sex, income, education, living status, alcohol intake, smoking status, physical activity level, & BMI
```{r}
Hannum_CESD_model_full<-lmer(CESD_10~Hannum_Age_COM_0*timefactor + (1 + time|ID) + Age + Sex + Income_Level + Education4 + Living_Status + 
                               Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_Full_Complete)
summary(Hannum_CESD_model_full)
anova(Hannum_CESD_model_full)#significant Hannum Age x time interaction
```


### 3.3.3: Simple Slopes Analyses (Contrasts)
```{r}
mean(COVID.Final$Hannum_Age_COM_0) #Mean Hannum Age = 63.00277
mean(COVID.Final$Hannum_Age_COM_0) + 2*sd(COVID.Final$Hannum_Age_COM_0) #High" Hannum Age = 87.13756 
mean(COVID.Final$Hannum_Age_COM_0) - 2*sd(COVID.Final$Hannum_Age_COM_0) #"Low" Hannum Age = 38.86798

ef2<-effect(term = "Hannum_Age_COM_0*timefactor", xlevels=list(Hannum_Age_COM_0=c(38.86798,63.00277,87.13756)), mod=Hannum_CESD_model_full)

ef2data<-as.data.frame(ef2)
ef2data2<-ef2data[c(2,3,4,1)]

ef2data2$Hannum<-NA
ef2data2$Hannum[ef2data2$Hannum_Age_COM_0 == 38.86798] <- "Low"
ef2data2$Hannum[ef2data2$Hannum_Age_COM_0 == 63.00277] <- "Mean"
ef2data2$Hannum[ef2data2$Hannum_Age_COM_0 == 87.13756] <- "High"

ef2data2

#Contrast 1: Differences between High and Low Hannum Age from BL to FU1
Diff4= ef2data2$fit[c(3)] - ef2data2$fit[c(6)] 
Diff5= ef2data2$fit[c(1)] - ef2data2$fit[c(4)]
Diff6= Diff4 - Diff5
se4 = sqrt((ef2data2$se[c(3)])^2 + (ef2data2$se[c(6)])^2)
se5 = sqrt((ef2data2$se[c(1)])^2 + (ef2data2$se[c(4)])^2)
se6 = sqrt(se4^2 + se5^2)
t3 = Diff6/se6
p3=pt(-abs(t3),df=n-1)
LL6= Diff6 - 1.64*se6
UL6=  Diff6 + 1.64*se6
Diff4
Diff5
Diff6
se4
se5
se6
t3
p3
LL6
UL6

#Contrast 2: Differences between High and Low Hannum Age from FU1 to COVID BL
Diff4.2= ef2data2$fit[c(6)] - ef2data2$fit[c(9)] 
Diff5.2= ef2data2$fit[c(4)] - ef2data2$fit[c(7)]
Diff6.2= Diff4.2 - Diff5.2
se4.2 = sqrt((ef2data2$se[c(6)])^2 + (ef2data2$se[c(9)])^2)
se5.2 = sqrt((ef2data2$se[c(4)])^2 + (ef2data2$se[c(7)])^2)
se6.2 = sqrt(se4.2^2 + se5.2^2)
t4 = Diff6.2/se6.2
p4=pt(-abs(t4),df=n-1)
LL6.2= Diff6.2 - 1.64*se6.2
UL6.2=  Diff6.2 + 1.64*se6.2
Diff4.2
Diff5.2
Diff6.2
se4.2
se5.2
se6.2
t4
p4
LL6.2
UL6.2

#Contrast 3: Differences between High and Low Hannum from COVID BL to COVID Exit
Diff4.3= ef2data2$fit[c(9)] - ef2data2$fit[c(12)] 
Diff5.3= ef2data2$fit[c(7)] - ef2data2$fit[c(10)]
Diff6.3= Diff4.3 - Diff5.3
se4.3 = sqrt((ef2data2$se[c(6)])^2 + (ef2data2$se[c(9)])^2)
se5.3 = sqrt((ef2data2$se[c(4)])^2 + (ef2data2$se[c(7)])^2)
se6.3 = sqrt(se4.3^2 + se5.3^2)
t3.3 = Diff6.3/se6.3
p3.3=pt(-abs(t3),df=n-1)
LL3.3= Diff6.3 - 1.64*se6.3
UL3.3=  Diff6.3 + 1.64*se6.3
Diff4.3
Diff5.3
Diff6.3
se4.3
se5.3
se6.3
t3.3
p3.3
LL3.3
UL3.3

```


### 3.3.4: Graphing Simple Slopes
```{r}
ef2data$Hannum<-NA
ef2data$Hannum[ef2data$Hannum_Age_COM_0==38.86798]<-"-2 SD"
ef2data$Hannum[ef2data$Hannum_Age_COM_0==63.00277]<-"Mean"
ef2data$Hannum[ef2data$Hannum_Age_COM_0==87.13756]<-"+2 SD"

ef2data$Timepoint<-NA
ef2data$Timepoint[ef2data$timefactor==1]<-"Baseline"
ef2data$Timepoint[ef2data$timefactor==2]<-"First Follow-up"
ef2data$Timepoint[ef2data$timefactor==3]<-"COVID-19 Study Entry"
ef2data$Timepoint[ef2data$timefactor==4]<-"COVID-19 Study Exit"
ef2data$Timepoint<-ordered(ef2data$Timepoint, levels = c("Baseline", "First Follow-up", "COVID-19 Study Entry", "COVID-19 Study Exit"))

ggplot(data=ef2data, aes(x=Timepoint, y=fit, group=Hannum)) + geom_line(aes(color=Hannum)) + ylim(0,10) + 
  labs(y= "CESD-10 Score", color="Hannum Age") + theme_classic() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                                                                         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```



## 3.4: Age Acceleration Difference


### 3.4.1: Crude Model
```{r}
AgeAccel_CESD_model_crude<-lmer(CESD_10~AgeAccelerationResidual_COM_0*timefactor + (1 + time|ID), REML = TRUE, COVID_Full_Complete)
summary(AgeAccel_CESD_model_crude)
anova(AgeAccel_CESD_model_crude)#no significant age accleration residual x time interaction
```


### 3.4.2: Full Model - Adjusted for age, sex, income, education, living status, alcohol intake, smoking status, physical activity level, & BMI
```{r}
AgeAccel_CESD_model_full<-lmer(CESD_10~AgeAccelerationResidual_COM_0*timefactor + (1 + time|ID) + Age + Sex + Income_Level + Education4 + Living_Status + 
                                 Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_Full_Complete)
summary(AgeAccel_CESD_model_full)
anova(AgeAccel_CESD_model_full)#no significant age accleration residual x time effect
```


### 3.4.3: Simple Slopes Analyses (Contrasts)
```{r}
mean(COVID.Final$AgeAccelerationResidual_COM_0) #Mean DNA Methylation Age = -0.15
mean(COVID.Final$AgeAccelerationResidual_COM_0) + 2*sd(COVID.Final$AgeAccelerationResidual_COM_0) #High" DNA Methylation Age = 8.73 
mean(COVID.Final$AgeAccelerationResidual_COM_0) - 2*sd(COVID.Final$AgeAccelerationResidual_COM_0) #"Low" DNA Methylation Age = -9.03


ef3<-effect(term = "AgeAccelerationResidual_COM_0*timefactor", xlevels=list(AgeAccelerationResidual_COM_0=c(-9.03,-0.15,8.73)), mod=AgeAccel_CESD_model_full)
efdata3<-as.data.frame(ef3)
efdata3<-efdata3[c(2,3,4,1)]

efdata3$DNA<-NA
efdata3$DNA[efdata3$AgeAccelerationResidual_COM_0 == -9.03] <- "Low"
efdata3$DNA[efdata3$AgeAccelerationResidual_COM_0 == -0.15] <- "Mean"
efdata3$DNA[efdata3$AgeAccelerationResidual_COM_0 == 8.73] <- "High"

efdata3

#Contrast 1: Differences between High and Low Age Acceleration Residual from BL to FU1
Diff7= efdata3$fit[c(3)] - efdata3$fit[c(6)] 
Diff8= efdata3$fit[c(1)] - efdata3$fit[c(4)]
Diff9= Diff7 - Diff8
se7 = sqrt((efdata3$se[c(3)])^2 + (efdata3$se[c(6)])^2)
se8 = sqrt((efdata3$se[c(1)])^2 + (efdata3$se[c(4)])^2)
se9 = sqrt(se7^2 + se8^2)
t9 = Diff9/se9
p9=pt(-abs(t9),df=n-1)
LL9= Diff9 - 1.64*se9
UL9=  Diff9 + 1.64*se9
Diff7
Diff8
Diff9
se7
se8
se9
t9
p9
LL9
UL9

#Contrast 2: Differences between High and Low Hannum Age from FU1 to COVID BL
Diff7.2= efdata3$fit[c(6)] - efdata3$fit[c(9)] 
Diff8.2= efdata3$fit[c(4)] - efdata3$fit[c(7)]
Diff9.2= Diff7.2 - Diff8.2
se7.2 = sqrt((efdata3$se[c(6)])^2 + (efdata3$se[c(9)])^2)
se8.2 = sqrt((efdata3$se[c(4)])^2 + (efdata3$se[c(7)])^2)
se9.2 = sqrt(se7.2^2 + se8.2^2)
t9.2 = Diff9.2/se9.2
p9.2=pt(-abs(t9.2),df=n-1)
LL9.2= Diff9.2 - 1.64*se9.2
UL9.2=  Diff9.2 + 1.64*se9.2
Diff7.2
Diff8.2
Diff9.2
se7.2
se8.2
se9.2
t9.2
p9.2
LL9.2
UL9.2

#Contrast 3: Differences between High and Low Hannum from COVID BL to COVID Exit
Diff7.3= efdata3$fit[c(9)] - efdata3$fit[c(12)] 
Diff8.3= efdata3$fit[c(7)] - efdata3$fit[c(10)]
Diff9.3= Diff7.3 - Diff8.3
se7.3 = sqrt((efdata3$se[c(6)])^2 + (efdata3$se[c(9)])^2)
se8.3 = sqrt((efdata3$se[c(4)])^2 + (efdata3$se[c(7)])^2)
se9.3 = sqrt(se7.3^2 + se8.3^2)
t9.3 = Diff9.3/se9.3
p9.3=pt(-abs(t9.3),df=n-1)
LL9.3= Diff9.3 - 1.64*se9.3
UL9.3=  Diff9.3 + 1.64*se9.3
Diff7.3
Diff8.3
Diff9.3
se7.3
se8.3
se9.3
t9.3
p9.3
LL9.3
UL9.3
```



# Step 4: Predicting Sex Differences in the Association of Epigenetic Age with Depressive Symptoms


## 4.1: DNA Methylation Age


### 4.1.1: Crude Model
```{r}
DNAbySex_CESD_model_crude<-lmer(CESD_10~DNAmAge_COM_0*timefactor*Sex + (1 + time|ID), REML = TRUE, COVID_Full_Complete)
summary(DNAbySex_CESD_model_crude)
anova(DNAbySex_CESD_model_crude)#significant sex x time interaction, but not sex x time x DNA methylation interaction
```


### 4.1.2: Full Model - Adjusted for age, income, education, living status, alcohol intake, smoking status, physical activity level, & BMI
```{r}
DNAbySex_CESD_model_full<-lmer(CESD_10~DNAmAge_COM_0*timefactor*Sex + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                                 Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_Full_Complete)
summary(DNAbySex_CESD_model_full)
anova(DNAbySex_CESD_model_full)#significant sex x time interaction, but not sex x time x DNA methylation interaction
```


### 4.1.3: Simple Slopes
```{r}
describeBy(COVID.Final.BL$DNAmAge_COM_0, COVID.Final.BL$Sex)
  #Males
58.77 + 2*8.96#high= 76.69
58.77 - 2*8.96#low= 40.85

  #Females
56 + 2*8.18#high= 72.36
56 - 2*8.18#low= 39.64


#Only Males Model
COVID_male_only<-subset(COVID_Full_Complete, Sex == "M")
DNAmale<-lmer(CESD_10~DNAmAge_COM_0*timefactor + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                                 Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_male_only)

efmale1<-effect(term = "DNAmAge_COM_0*timefactor", xlevels=list(DNAmAge_COM_0=c(40.85, 76.69)), mod=DNAmale)

maledata1<-as.data.frame(efmale1)
maledata1<-maledata1[c(2,3,4,1)]


maledata1$DNA<-NA
maledata1$DNA[maledata1$DNAmAge_COM_0 == 40.85] <- "Low"
maledata1$DNA[maledata1$DNAmAge_COM_0 == 76.69] <- "High"

maledata1

#Contrast 1: Differences between High and Low Age Acceleration Residual from BL to FU1
Diff7= maledata1$fit[c(2)] - maledata1$fit[c(4)] 
Diff8= maledata1$fit[c(1)] - maledata1$fit[c(3)]
Diff9= Diff7 - Diff8
se7 = sqrt((maledata1$se[c(2)])^2 + (maledata1$se[c(4)])^2)
se8 = sqrt((maledata1$se[c(1)])^2 + (maledata1$se[c(3)])^2)
se9 = sqrt(se7^2 + se8^2)
t9 = Diff9/se9
p9=pt(-abs(t9),df=n-1)
LL9= Diff9 - 1.64*se9
UL9=  Diff9 + 1.64*se9
Diff7
Diff8
Diff9
se7
se8
se9
t9
p9
LL9
UL9

#Contrast 2: Differences between High and Low Hannum Age from FU1 to COVID BL
Diff7.2= maledata1$fit[c(4)] - maledata1$fit[c(6)] 
Diff8.2= maledata1$fit[c(3)] - maledata1$fit[c(5)]
Diff9.2= Diff7.2 - Diff8.2
se7.2 = sqrt((maledata1$se[c(4)])^2 + (maledata1$se[c(6)])^2)
se8.2 = sqrt((maledata1$se[c(3)])^2 + (maledata1$se[c(5)])^2)
se9.2 = sqrt(se7.2^2 + se8.2^2)
t9.2 = Diff9.2/se9.2
p9.2=pt(-abs(t9.2),df=n-1)
LL9.2= Diff9.2 - 1.64*se9.2
UL9.2=  Diff9.2 + 1.64*se9.2
Diff7.2
Diff8.2
Diff9.2
se7.2
se8.2
se9.2
t9.2
p9.2
LL9.2
UL9.2

#Contrast 3: Differences between High and Low Hannum from COVID BL to COVID Exit
Diff7.3= maledata1$fit[c(6)] - maledata1$fit[c(8)] 
Diff8.3= maledata1$fit[c(5)] - maledata1$fit[c(7)]
Diff9.3= Diff7.3 - Diff8.3
se7.3 = sqrt((maledata1$se[c(6)])^2 + (maledata1$se[c(8)])^2)
se8.3 = sqrt((maledata1$se[c(5)])^2 + (maledata1$se[c(7)])^2)
se9.3 = sqrt(se7.3^2 + se8.3^2)
t9.3 = Diff9.3/se9.3
p9.3=pt(-abs(t9.3),df=n-1)
LL9.3= Diff9.3 - 1.64*se9.3
UL9.3=  Diff9.3 + 1.64*se9.3
Diff7.3
Diff8.3
Diff9.3
se7.3
se8.3
se9.3
t9.3
p9.3
LL9.3
UL9.3


#Only Females Model
COVID_female_only<-subset(COVID_Full_Complete, Sex == "F")
DNAfemale<-lmer(CESD_10~DNAmAge_COM_0*timefactor + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_female_only)

female1<-effect(term = "DNAmAge_COM_0*timefactor", xlevels=list(DNAmAge_COM_0=c(39.64, 72.36)), mod=DNAfemale)

femaledata1<-as.data.frame(female1)
femaledata1<-femaledata1[c(2,3,4,1)]


femaledata1$DNA<-NA
femaledata1$DNA[femaledata1$DNAmAge_COM_0 == 40.85] <- "Low"
femaledata1$DNA[femaledata1$DNAmAge_COM_0 == 76.69] <- "High"

femaledata1

#Contrast 1: Differences between High and Low Age Acceleration Residual from BL to FU1
Diff7= femaledata1$fit[c(2)] - femaledata1$fit[c(4)] 
Diff8= femaledata1$fit[c(1)] - femaledata1$fit[c(3)]
Diff9= Diff7 - Diff8
se7 = sqrt((femaledata1$se[c(2)])^2 + (femaledata1$se[c(4)])^2)
se8 = sqrt((femaledata1$se[c(1)])^2 + (femaledata1$se[c(3)])^2)
se9 = sqrt(se7^2 + se8^2)
t9 = Diff9/se9
p9=pt(-abs(t9),df=n-1)
LL9= Diff9 - 1.64*se9
UL9=  Diff9 + 1.64*se9
Diff7
Diff8
Diff9
se7
se8
se9
t9
p9
LL9
UL9

#Contrast 2: Differences between High and Low Hannum Age from FU1 to COVID BL
Diff7.2= femaledata1$fit[c(4)] - femaledata1$fit[c(6)] 
Diff8.2= femaledata1$fit[c(3)] - femaledata1$fit[c(5)]
Diff9.2= Diff7.2 - Diff8.2
se7.2 = sqrt((femaledata1$se[c(4)])^2 + (femaledata1$se[c(6)])^2)
se8.2 = sqrt((femaledata1$se[c(3)])^2 + (femaledata1$se[c(5)])^2)
se9.2 = sqrt(se7.2^2 + se8.2^2)
t9.2 = Diff9.2/se9.2
p9.2=pt(-abs(t9.2),df=n-1)
LL9.2= Diff9.2 - 1.64*se9.2
UL9.2=  Diff9.2 + 1.64*se9.2
Diff7.2
Diff8.2
Diff9.2
se7.2
se8.2
se9.2
t9.2
p9.2
LL9.2
UL9.2

#Contrast 3: Differences between High and Low Hannum from COVID BL to COVID Exit
Diff7.3= femaledata1$fit[c(6)] - femaledata1$fit[c(8)] 
Diff8.3= femaledata1$fit[c(5)] - femaledata1$fit[c(7)]
Diff9.3= Diff7.3 - Diff8.3
se7.3 = sqrt((femaledata1$se[c(6)])^2 + (femaledata1$se[c(8)])^2)
se8.3 = sqrt((femaledata1$se[c(5)])^2 + (femaledata1$se[c(7)])^2)
se9.3 = sqrt(se7.3^2 + se8.3^2)
t9.3 = Diff9.3/se9.3
p9.3=pt(-abs(t9.3),df=n-1)
LL9.3= Diff9.3 - 1.64*se9.3
UL9.3=  Diff9.3 + 1.64*se9.3
Diff7.3
Diff8.3
Diff9.3
se7.3
se8.3
se9.3
t9.3
p9.3
LL9.3
UL9.3
```


## 4.2: Hannum Age


### 4.2.1: Crude Model
```{r}
HannumbySex_CESD_model_crude<-lmer(CESD_10~Hannum_Age_COM_0*timefactor*Sex + (1 + time|ID), REML = TRUE, COVID_Full_Complete)
summary(HannumbySex_CESD_model_crude)
anova(HannumbySex_CESD_model_crude)#no significant sex interaction effects
```


### 4.2.2: Full Model - Adjusted for age, income, education, living status, alcohol intake, smoking status, physical activity level, & BMI
```{r}
HannumbySex_CESD_model_full<-lmer(CESD_10~Hannum_Age_COM_0*timefactor*Sex + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                                    Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_Full_Complete)
summary(HannumbySex_CESD_model_full)
anova(HannumbySex_CESD_model_full)#no significant sex interaction effects
```


### 4.2.3: Simple Slopes
```{r}
describeBy(COVID.Final.BL$Hannum_Age_COM_0, COVID.Final.BL$Sex)
#Males
64.97 + 2*11.85#high= 88.67
64.97 - 2*11.85#low= 41.27

#Females
60.69 + 2*11.92#high= 84.53
60.69 - 2*11.92#low= 36.85


#Only Males Model
DNAmale<-lmer(CESD_10~Hannum_Age_COM_0*timefactor + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_male_only)

efmale1<-effect(term = "Hannum_Age_COM_0*timefactor", xlevels=list(Hannum_Age_COM_0=c(41.27, 88.67)), mod=DNAmale)

maledata1<-as.data.frame(efmale1)
maledata1<-maledata1[c(2,3,4,1)]


maledata1$DNA<-NA
maledata1$DNA[maledata1$Hannum_Age_COM_0 == 41.27] <- "Low"
maledata1$DNA[maledata1$Hannum_Age_COM_0 == 88.67] <- "High"

maledata1

#Contrast 1: Differences between High and Low Age Acceleration Residual from BL to FU1
Diff7= maledata1$fit[c(2)] - maledata1$fit[c(4)] 
Diff8= maledata1$fit[c(1)] - maledata1$fit[c(3)]
Diff9= Diff7 - Diff8
se7 = sqrt((maledata1$se[c(2)])^2 + (maledata1$se[c(4)])^2)
se8 = sqrt((maledata1$se[c(1)])^2 + (maledata1$se[c(3)])^2)
se9 = sqrt(se7^2 + se8^2)
t9 = Diff9/se9
p9=pt(-abs(t9),df=n-1)
LL9= Diff9 - 1.64*se9
UL9=  Diff9 + 1.64*se9
Diff7
Diff8
Diff9
se7
se8
se9
t9
p9
LL9
UL9

#Contrast 2: Differences between High and Low Hannum Age from FU1 to COVID BL
Diff7.2= maledata1$fit[c(4)] - maledata1$fit[c(6)] 
Diff8.2= maledata1$fit[c(3)] - maledata1$fit[c(5)]
Diff9.2= Diff7.2 - Diff8.2
se7.2 = sqrt((maledata1$se[c(4)])^2 + (maledata1$se[c(6)])^2)
se8.2 = sqrt((maledata1$se[c(3)])^2 + (maledata1$se[c(5)])^2)
se9.2 = sqrt(se7.2^2 + se8.2^2)
t9.2 = Diff9.2/se9.2
p9.2=pt(-abs(t9.2),df=n-1)
LL9.2= Diff9.2 - 1.64*se9.2
UL9.2=  Diff9.2 + 1.64*se9.2
Diff7.2
Diff8.2
Diff9.2
se7.2
se8.2
se9.2
t9.2
p9.2
LL9.2
UL9.2

#Contrast 3: Differences between High and Low Hannum from COVID BL to COVID Exit
Diff7.3= maledata1$fit[c(6)] - maledata1$fit[c(8)] 
Diff8.3= maledata1$fit[c(5)] - maledata1$fit[c(7)]
Diff9.3= Diff7.3 - Diff8.3
se7.3 = sqrt((maledata1$se[c(6)])^2 + (maledata1$se[c(8)])^2)
se8.3 = sqrt((maledata1$se[c(5)])^2 + (maledata1$se[c(7)])^2)
se9.3 = sqrt(se7.3^2 + se8.3^2)
t9.3 = Diff9.3/se9.3
p9.3=pt(-abs(t9.3),df=n-1)
LL9.3= Diff9.3 - 1.64*se9.3
UL9.3=  Diff9.3 + 1.64*se9.3
Diff7.3
Diff8.3
Diff9.3
se7.3
se8.3
se9.3
t9.3
p9.3
LL9.3
UL9.3


#Only Females Model
DNAfemale<-lmer(CESD_10~Hannum_Age_COM_0*timefactor + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                  Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_female_only)

female1<-effect(term = "Hannum_Age_COM_0*timefactor", xlevels=list(Hannum_Age_COM_0=c(36.85, 84.53)), mod=DNAfemale)

femaledata1<-as.data.frame(female1)
femaledata1<-femaledata1[c(2,3,4,1)]


femaledata1$DNA<-NA
femaledata1$DNA[femaledata1$Hannum_Age_COM_0 == 36.85] <- "Low"
femaledata1$DNA[femaledata1$Hannum_Age_COM_0 == 84.53] <- "High"

femaledata1

#Contrast 1: Differences between High and Low Age Acceleration Residual from BL to FU1
Diff7= femaledata1$fit[c(2)] - femaledata1$fit[c(4)] 
Diff8= femaledata1$fit[c(1)] - femaledata1$fit[c(3)]
Diff9= Diff7 - Diff8
se7 = sqrt((femaledata1$se[c(2)])^2 + (femaledata1$se[c(4)])^2)
se8 = sqrt((femaledata1$se[c(1)])^2 + (femaledata1$se[c(3)])^2)
se9 = sqrt(se7^2 + se8^2)
t9 = Diff9/se9
p9=pt(-abs(t9),df=n-1)
LL9= Diff9 - 1.64*se9
UL9=  Diff9 + 1.64*se9
Diff7
Diff8
Diff9
se7
se8
se9
t9
p9
LL9
UL9

#Contrast 2: Differences between High and Low Hannum Age from FU1 to COVID BL
Diff7.2= femaledata1$fit[c(4)] - femaledata1$fit[c(6)] 
Diff8.2= femaledata1$fit[c(3)] - femaledata1$fit[c(5)]
Diff9.2= Diff7.2 - Diff8.2
se7.2 = sqrt((femaledata1$se[c(4)])^2 + (femaledata1$se[c(6)])^2)
se8.2 = sqrt((femaledata1$se[c(3)])^2 + (femaledata1$se[c(5)])^2)
se9.2 = sqrt(se7.2^2 + se8.2^2)
t9.2 = Diff9.2/se9.2
p9.2=pt(-abs(t9.2),df=n-1)
LL9.2= Diff9.2 - 1.64*se9.2
UL9.2=  Diff9.2 + 1.64*se9.2
Diff7.2
Diff8.2
Diff9.2
se7.2
se8.2
se9.2
t9.2
p9.2
LL9.2
UL9.2

#Contrast 3: Differences between High and Low Hannum from COVID BL to COVID Exit
Diff7.3= femaledata1$fit[c(6)] - femaledata1$fit[c(8)] 
Diff8.3= femaledata1$fit[c(5)] - femaledata1$fit[c(7)]
Diff9.3= Diff7.3 - Diff8.3
se7.3 = sqrt((femaledata1$se[c(6)])^2 + (femaledata1$se[c(8)])^2)
se8.3 = sqrt((femaledata1$se[c(5)])^2 + (femaledata1$se[c(7)])^2)
se9.3 = sqrt(se7.3^2 + se8.3^2)
t9.3 = Diff9.3/se9.3
p9.3=pt(-abs(t9.3),df=n-1)
LL9.3= Diff9.3 - 1.64*se9.3
UL9.3=  Diff9.3 + 1.64*se9.3
Diff7.3
Diff8.3
Diff9.3
se7.3
se8.3
se9.3
t9.3
p9.3
LL9.3
UL9.3
```



## 4.3: Age Acceleration Difference


### 4.3.1: Crude Model
```{r}
AgeAccelbySex_CESD_model_crude<-lmer(CESD_10~AgeAccelerationResidual_COM_0*timefactor*Sex + (1 + time|ID), REML = TRUE, COVID_Full_Complete)
summary(AgeAccelbySex_CESD_model_crude)
anova(AgeAccelbySex_CESD_model_crude)#significant sex by time effect, but no sex by time by age acceleration difference interaction
```


### 4.3.2: Full Model - Adjusted for age, sex, income, education, living status, alcohol intake, smoking status, physical activity level, & BMI
```{r}
AgeAccelbySex_CESD_model_full<-lmer(CESD_10~AgeAccelerationResidual_COM_0*timefactor*Sex + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                                      Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_Full_Complete)
summary(AgeAccelbySex_CESD_model_full)
anova(AgeAccelbySex_CESD_model_full)#no significant sex interaction effects
```


### 4.3.3: Simple Slopes
```{r}
describeBy(COVID.Final.BL$AgeAccelerationResidual_COM_0, COVID.Final.BL$Sex)
#Males
0.81 + 2*4.6#high= 10.01
0.81 - 2*4.6#low= -8.39

#Females
-1.27 + 2*3.96#high= 6.65
-1.27 - 2*3.96#low= -9.19


#Only Males Model
DNAmale<-lmer(CESD_10~AgeAccelerationResidual_COM_0*timefactor + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_male_only)

efmale1<-effect(term = "AgeAccelerationResidual_COM_0*timefactor", xlevels=list(AgeAccelerationResidual_COM_0=c(-8.39, 10.01)), mod=DNAmale)

maledata1<-as.data.frame(efmale1)
maledata1<-maledata1[c(2,3,4,1)]


maledata1$DNA<-NA
maledata1$DNA[maledata1$AgeAccelerationResidual_COM_0 == -8.39] <- "Low"
maledata1$DNA[maledata1$AgeAccelerationResidual_COM_0 == 10.01] <- "High"

maledata1

#Contrast 1: Differences between High and Low Age Acceleration Residual from BL to FU1
Diff7= maledata1$fit[c(2)] - maledata1$fit[c(4)] 
Diff8= maledata1$fit[c(1)] - maledata1$fit[c(3)]
Diff9= Diff7 - Diff8
se7 = sqrt((maledata1$se[c(2)])^2 + (maledata1$se[c(4)])^2)
se8 = sqrt((maledata1$se[c(1)])^2 + (maledata1$se[c(3)])^2)
se9 = sqrt(se7^2 + se8^2)
t9 = Diff9/se9
p9=pt(-abs(t9),df=n-1)
LL9= Diff9 - 1.64*se9
UL9=  Diff9 + 1.64*se9
Diff7
Diff8
Diff9
se7
se8
se9
t9
p9
LL9
UL9

#Contrast 2: Differences between High and Low Hannum Age from FU1 to COVID BL
Diff7.2= maledata1$fit[c(4)] - maledata1$fit[c(6)] 
Diff8.2= maledata1$fit[c(3)] - maledata1$fit[c(5)]
Diff9.2= Diff7.2 - Diff8.2
se7.2 = sqrt((maledata1$se[c(4)])^2 + (maledata1$se[c(6)])^2)
se8.2 = sqrt((maledata1$se[c(3)])^2 + (maledata1$se[c(5)])^2)
se9.2 = sqrt(se7.2^2 + se8.2^2)
t9.2 = Diff9.2/se9.2
p9.2=pt(-abs(t9.2),df=n-1)
LL9.2= Diff9.2 - 1.64*se9.2
UL9.2=  Diff9.2 + 1.64*se9.2
Diff7.2
Diff8.2
Diff9.2
se7.2
se8.2
se9.2
t9.2
p9.2
LL9.2
UL9.2

#Contrast 3: Differences between High and Low Hannum from COVID BL to COVID Exit
Diff7.3= maledata1$fit[c(6)] - maledata1$fit[c(8)] 
Diff8.3= maledata1$fit[c(5)] - maledata1$fit[c(7)]
Diff9.3= Diff7.3 - Diff8.3
se7.3 = sqrt((maledata1$se[c(6)])^2 + (maledata1$se[c(8)])^2)
se8.3 = sqrt((maledata1$se[c(5)])^2 + (maledata1$se[c(7)])^2)
se9.3 = sqrt(se7.3^2 + se8.3^2)
t9.3 = Diff9.3/se9.3
p9.3=pt(-abs(t9.3),df=n-1)
LL9.3= Diff9.3 - 1.64*se9.3
UL9.3=  Diff9.3 + 1.64*se9.3
Diff7.3
Diff8.3
Diff9.3
se7.3
se8.3
se9.3
t9.3
p9.3
LL9.3
UL9.3


#Only Females Model
DNAfemale<-lmer(CESD_10~AgeAccelerationResidual_COM_0*timefactor + (1 + time|ID) + Age + Income_Level + Education4 + Living_Status + 
                  Alcohol + Smoking_Status + PASE_TOTAL_0 + BMI_0, COVID_female_only)

female1<-effect(term = "AgeAccelerationResidual_COM_0*timefactor", xlevels=list(AgeAccelerationResidual_COM_0=c(-9.19, 6.65)), mod=DNAfemale)

femaledata1<-as.data.frame(female1)
femaledata1<-femaledata1[c(2,3,4,1)]


femaledata1$DNA<-NA
femaledata1$DNA[femaledata1$Hannum_Age_COM_0 == -9.19] <- "Low"
femaledata1$DNA[femaledata1$Hannum_Age_COM_0 == 6.65] <- "High"

femaledata1

#Contrast 1: Differences between High and Low Age Acceleration Residual from BL to FU1
Diff7= femaledata1$fit[c(2)] - femaledata1$fit[c(4)] 
Diff8= femaledata1$fit[c(1)] - femaledata1$fit[c(3)]
Diff9= Diff7 - Diff8
se7 = sqrt((femaledata1$se[c(2)])^2 + (femaledata1$se[c(4)])^2)
se8 = sqrt((femaledata1$se[c(1)])^2 + (femaledata1$se[c(3)])^2)
se9 = sqrt(se7^2 + se8^2)
t9 = Diff9/se9
p9=pt(-abs(t9),df=n-1)
LL9= Diff9 - 1.64*se9
UL9=  Diff9 + 1.64*se9
Diff7
Diff8
Diff9
se7
se8
se9
t9
p9
LL9
UL9

#Contrast 2: Differences between High and Low Hannum Age from FU1 to COVID BL
Diff7.2= femaledata1$fit[c(4)] - femaledata1$fit[c(6)] 
Diff8.2= femaledata1$fit[c(3)] - femaledata1$fit[c(5)]
Diff9.2= Diff7.2 - Diff8.2
se7.2 = sqrt((femaledata1$se[c(4)])^2 + (femaledata1$se[c(6)])^2)
se8.2 = sqrt((femaledata1$se[c(3)])^2 + (femaledata1$se[c(5)])^2)
se9.2 = sqrt(se7.2^2 + se8.2^2)
t9.2 = Diff9.2/se9.2
p9.2=pt(-abs(t9.2),df=n-1)
LL9.2= Diff9.2 - 1.64*se9.2
UL9.2=  Diff9.2 + 1.64*se9.2
Diff7.2
Diff8.2
Diff9.2
se7.2
se8.2
se9.2
t9.2
p9.2
LL9.2
UL9.2

#Contrast 3: Differences between High and Low Hannum from COVID BL to COVID Exit
Diff7.3= femaledata1$fit[c(6)] - femaledata1$fit[c(8)] 
Diff8.3= femaledata1$fit[c(5)] - femaledata1$fit[c(7)]
Diff9.3= Diff7.3 - Diff8.3
se7.3 = sqrt((femaledata1$se[c(6)])^2 + (femaledata1$se[c(8)])^2)
se8.3 = sqrt((femaledata1$se[c(5)])^2 + (femaledata1$se[c(7)])^2)
se9.3 = sqrt(se7.3^2 + se8.3^2)
t9.3 = Diff9.3/se9.3
p9.3=pt(-abs(t9.3),df=n-1)
LL9.3= Diff9.3 - 1.64*se9.3
UL9.3=  Diff9.3 + 1.64*se9.3
Diff7.3
Diff8.3
Diff9.3
se7.3
se8.3
se9.3
t9.3
p9.3
LL9.3
UL9.3
